---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.1'
      jupytext_version: 1.2.4
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# R by Example

Alex Chi

```{r}
library(repr)
options(repr.plot.width=10, repr.plot.height=8)
```

## Chapter 3 Exercises

## 3.1 (Fast food eating preference).

Fifteen students in a statistics class were asked to state their preference among the three restaurtants Wendys, McDonalds, and Subway. The responses for the students are presented below.

```
Wendys McDonalds Subway Subway Subway Wendys 
Wendys Subway Wendys Subway Subway Subway
Subway Subway Subway
```
a. Use the scan function to read these data into the R command window.

b. Use the table function to find the frequencies of students who prefer the three restaurants.

c. Compute the proportions of students in each category.

d. Construct two different graphical displays of the proportions.

```{r}
# scan cannot be used in Jupyter R, so I just construct the table
restaurants = unlist(strsplit("Wendys McDonalds Subway Subway Subway Wendys Wendys Subway Wendys Subway Subway Subway Subway Subway Subway", split=' '))
freq = table(factor(restaurants))
freq
```

```{r}
freq / length(restaurants)
```

```{r}
barplot(prop.table(freq))
```

## 3.2 (Dice rolls)
Suppose you roll a pair of dice 1000 times
a. One can simulate 1000 rolls of a fair die using the R function sample(6, 1000, replace=TRUE). Using this function twice, store 1000 simulated rolls of the first die in the variable die1 and 1000 simulated rolls of the second die in the variable die2.

b. For each pair of rolls, compute the sum of rolls, and store the sums in the variable die.sum.

c. Use the table function to tabulate the values of the sum of die rolls. Com- pute the proportions for each sum value and compare these proportions with the exact probabilities of the sum of two die rolls

```{r}
die1 = sample(6, 1000, replace=TRUE)
die2 = sample(6, 1000, replace=TRUE)
die.sum = die1 + die2
table(die.sum) / length(die.sum)
```

```{r}
k = 2:12
die.expected = (6 - abs(k - (6 + 1))) / (6 * 6)
die.freq = table(die.sum) / length(die.sum)
plot(k, die.freq, type="h", lwd=2, lty=1, ylab="Freq")
lines(k + .2, die.expected, type="h", lwd=2, lty=2) 
```

## 3.3 (Does baseball hitting data follow a binomial distribution?)

Albert Pujols is a baseball player who has n opportunities to hit in a single game. If y denotes the number of hits for a game, then it is reasonable to assume that y has a binomial distribution with sample size n and probability of success p = 0.312, where 0.312 is Pujols’ batting average (success rate) for the 2010 baseball season.

a. In 70 games Pujols had exactly n = 4 opportunities to hit and the num- ber of hits y in these 70 games is tabulated in the following table. Use the dbinom function to compute the expected counts and the chisq.test function to test if the counts follow a binomial(4, 0.312) distribution.

b. In 25 games Pujols had exactly n = 5 opportunities to hit and the number of hits y in these 25 games is shown in the table below. Use the chisq.test function to test if the counts follow a binomial(5, 0.312) distribution

```{r}
n = 4
k = 0 : (n - 1)
baseball.expected = dbinom(k, n, 0.312)
baseball.game = c(17, 31, 17, 5)
baseball.game = baseball.game / 70
names(baseball.game) = k
plot(k, baseball.game, type="h", lwd=2, lty=1, ylab="Freq")
lines(k + .2, baseball.expected, type="h", lwd=2, lty=2) 
```

```{r}
baseball.expected = dbinom(k, 5, 0.312)
baseball.game = c(5, 5, 4, 11)
chisq.test(baseball.game, p=baseball.expected)
```

## 3.4 (Categorizing ages in the twins dataset). The variable AGE gives the age (in years) of twin 1.
a. Use the cut function on AGE with the breakpoints 30, 40, and 50 to create a categorized version of the twin’s age.

b. Use the table function to find the frequencies in the four age categories.

c. Construct a graph of the proportions in the four age categories

```{r}
data(twins)
```

## 3.8 (Are the digits of π random?). 
The National Institute of Standards and Technology has a web page that lists the first 5000 digits of the irrational number π. One can read these digits into R by means of the script

```{r}
pidigits = read.table("http://www.itl.nist.gov/div898/strd/univ/data/PiDigits.dat", skip=60)
```

```{r}
barplot(prop.table(table(pidigits)))
```

```{r}
chisq.test(table(pidigits), punif(1:10))
```

```{r}

```
